project scope:



Library the packages that might be used later
```{r}
library(tidyverse)
library(sf)
library(tmap)
library(janitor)
library(spatstat)
library(spdep)
library(RColorBrewer)
library(tmaptools)
library(here)
```

Loading the data
```{r}
#drop the NAs to prevent errors when plotting
graffiti <- read_csv(here("Graffiti.csv")) %>%
  na.omit()
shape <- st_read(here("Census 2010 Tracts", "geo_export_fd7daba4-f63c-4100-8878-9512e08d03fc.shp"))
```
Reviewing the variable types to ensure that there are no columns represented as character types that should be numeric due to the presence of missing values
```{r}
Datatypelist <- graffiti %>%
  summarise_all(class) %>%
  pivot_longer(everything(),
               names_to="All_variables",
               values_to="Variable_class")
Datatypelist
```
As the point column shows, the location is being recorded in latitude & longitude. We need to first convert them into sf object and set the lcoal crs https://epsg.io/7131
```{r}
points <- graffiti %>%
  mutate(
    x = as.numeric(gsub("\\((.*),.*\\)", "\\1", Point)),
    y = as.numeric(gsub("\\(.*, (.*)\\)", "\\1", Point))
  ) %>%
  #remove the remaining NA values
  filter(x>0 & y<0) %>%
  #standardise the column names
  clean_names() %>% 
  st_as_sf(., coords = c("y", "x"), 
                   crs = 4326) 
```
project shape object into local crs
```{r}
shape_projected <- shape %>%
  st_transform(., 7131) 

shape_filtered_projected <- shape_projected[-190, ]
```
project the points
```{r}
projected_points <- points %>%
  st_transform(., crs=7131)
```



quickly plot the data to get an overall sense and potential distribution of the data
```{r}
tmap_mode("plot")
tm_shape(shape) +
  tm_polygons(col = NA, alpha = 0.5) +
tm_shape(points) +
  tm_dots(col = "blue")
```
spatial subset
```{r}
graffiti_within <- projected_points[shape_projected, ,op=st_intersects]

```

select offensive ones
```{r}
points_2023 <- projected_points %>% 
  filter(str_detect(opened, "2023")) %>%
  filter(!str_detect(status_notes, "duplicate")) %>% 
  filter(!str_detect(request_details, "Not_Offensive")) 


```

create a window
```{r}
window <- as.owin(shape_projected)
plot(window)
```

```{r}
#create a sp object
points_2023_sp<- points_2023 %>%
  as(., 'Spatial')
#create a ppp object
points_2023_sp.ppp <- ppp(x=points_2023_sp@coords[,1],
                          y=points_2023_sp@coords[,2],
                          window=window)
```


plot the ppp
```{r}
points_2023_sp.ppp %>%
  plot(.,pch=16,cex=0.5, 
       main="San Francisco Graffiti 2023")
```
Ripley's K
```{r}
K <- points_2023_sp.ppp %>%
  Kest(., correction="border") %>%
  plot()
```
not useful

```{r}
library(raster)
library(fpc)
library(dbscan)
library(ggplot2)

#first extract the points from the spatial points data frame
points_2023_extracted <- points_2023_sp %>%
coordinates(.)%>%
as.data.frame()

#now run the dbscan analysis
db <- points_2023_extracted %>%
fpc::dbscan(.,eps = 200, MinPts = 500)

#now plot the results
plot(db, points_2023_extracted, main = "DBSCAN Output", frame = F)
plot(shape_projected$geometry, add=T)
```
```{r}
points_2023_extracted%>%
  dbscan::kNNdistplot(.,k=500)
points_2023_extracted <- points_2023_extracted %>%
  mutate(dbcluster=db$cluster)
```
```{r}
tosf <- points_2023_extracted%>%
  st_as_sf(., coords = c("coords.x1", "coords.x2"), 
                   crs = 7131)%>%
  filter(dbcluster>0)

ggplot(data = shape_projected) +
  # add the geometry of the community areas
  geom_sf() +
  # add the geometry of the points - i have had to set the data here to add the layer
  geom_sf(data = tosf, size = 0.4, colour=tosf$dbcluster, fill=tosf$dbcluster)
```

_____________________________________
Regression Model Atttempt

First, calculate the density 
```{r}
points_sf_joined <- shape_projected %>% 
  mutate(n = lengths(st_intersects(., points_2023)))%>%
  janitor::clean_names()%>%
  #calculate area
  mutate(area=st_area(.))%>%
  #then density of the points per ward
  mutate(density=n/area)
  #select density and some other variables 

tm_shape(points_sf_joined) +
    tm_polygons("density",
        style="jenks",
        palette="PuOr",
        title="Graffiti density")
```


```{r}
library(spdep)
#calculate the centroids


coordsW <- points_sf_joined%>%
  st_centroid()%>%
  st_geometry()
  
plot(coordsW,axes=TRUE)
```
```{r}
#create a neighbours list
Graffiti_nb <- points_sf_joined %>%
  poly2nb(., queen=T)
summary(Graffiti_nb)
```

```{r}
#plot them
plot(Graffiti_nb, st_geometry(coordsW), col="red")
#add a map underneath
plot(points_sf_joined$geometry, add=T)
```

```{r}
# make weight matrix
Graffiti_nb.lw <- Graffiti_nb %>%
  nb2mat(., style="W", zero.policy = TRUE)

sum(Graffiti_nb.lw)

# make weight list for Moran's I

Graffiti_nb.lw <- Graffiti_nb %>%
  nb2listw(., style="W", zero.policy = TRUE)
```
```{r}
Graffiti_Density <- points_sf_joined %>%
  pull(density) %>%
  as.vector()%>%
  moran.test(., Graffiti_nb.lw)

Graffiti_Density
```
```{r}
Graffiti_Density <- points_sf_joined %>%
  pull(density) %>%
  as.vector()%>%
  localmoran(., Graffiti_nb.lw)%>%
  as_tibble()

points_sf_joined <- points_sf_joined %>%
  mutate(density_I =as.numeric(Graffiti_Density$Ii))%>%
  mutate(density_Iz =as.numeric(Graffiti_Density$Z.Ii))
breaks1<-c(-1000,-2.58,-1.96,-1.65,1.65,1.96,2.58,1000)

library(RColorBrewer)
MoranColours<- rev(brewer.pal(8, "RdGy"))

tm_shape(points_sf_joined) +
    tm_polygons("density_Iz",
        style="fixed",
        breaks=breaks1,
        palette=MoranColours,
        midpoint=NA,
        title="Local Moran's I, Graffiti in San Francisco 2023")



```


-----------------------------
Regression


```{r}
library(tidycensus)
options(tigris_use_cache = TRUE)
v20 <- load_variables(2020, "acs5")
view(v20)


census_SF <- get_acs(
  state = "CA",
  county = "San Francisco",
  geography = "tract",
  variables = c("B27001_001", "B01003_001"),
  geometry = TRUE,
  year = 2020
)%>%
    st_transform(., crs=7131)
```

calculate the percentage of having health care
```{r}
health_percentage <- census_SF %>%
  filter(variable %in% c("B01003_001", "B27001_001")) %>%
  group_by(GEOID) %>%
  summarise(
    total_population = sum(estimate[variable == "B01003_001"]),  # Calculate total population
    health_insurance = sum(estimate[variable == "B27001_001"]),  # Calculate population with health insurance
    percentage = (health_insurance / total_population) * 100  # Calculate percentage
  ) %>%
  arrange(GEOID) %>% 
  filter(percentage != 0)


```

```{r}
tm_shape(health_percentage) +
    tm_polygons("percentage",
        style="jenks",
        palette="PuOr",
        title="health insurance density")
```

Scatter plot
```{r}
#drop geomatry
health_pure <- st_drop_geometry(health_percentage)
density_pure <- st_drop_geometry(points_sf_joined)

linear_san <- left_join(health_pure, density_pure, by = c("GEOID" = "geoid10")) %>% 
  na.omit()

library(units)
q <- qplot(x = `percentage`, 
           y = `density`, 
           data=linear_san)

#plot with a regression line - note, I've added some jitter here as the x-scale is rounded
q + stat_smooth(method="lm", se=FALSE, size=1) + 
  geom_jitter()
```




