project scope:



Library the packages that might be used later
```{r}
library(tidyverse)
library(sf)
library(tmap)
library(janitor)
library(spatstat)
library(spdep)
library(RColorBrewer)
library(tmaptools)
library(here)
```

Loading the data
```{r}
#drop the NAs to prevent errors when plotting
graffiti <- read_csv(here("Graffiti.csv")) %>%
  na.omit()
shape <- read_csv(here("Census_2010_Tracts_20231213.csv"))
```
Reviewing the variable types to ensure that there are no columns represented as character types that should be numeric due to the presence of missing values
```{r}
Datatypelist <- graffiti %>%
  summarise_all(class) %>%
  pivot_longer(everything(),
               names_to="All_variables",
               values_to="Variable_class")
Datatypelist
```
As the point column shows, the location is being recorded in latitude & longitude. We need to first convert them into sf object and set the standard global crs.
```{r}
points <- graffiti %>%
  mutate(
    x = as.numeric(gsub("\\((.*),.*\\)", "\\1", Point)),
    y = as.numeric(gsub("\\(.*, (.*)\\)", "\\1", Point))
  ) %>%
  #remove the remaining NA values
  filter(x>0 & y<0) %>%
  #standardise the column names
  clean_names() %>% 
  st_as_sf(., coords = c("x", "y"), 
                   crs = 7131) %>% 
   select(-point)


```




quickly plot the data to get an overall sense and potential distribution of the data
```{r}
tmap_mode("plot")
tm_shape(points) +
  tm_polygons(col = NA, alpha = 0.5) +
tm_shape(points) +
  tm_dots(col = "blue")
```





